Main Analyses of the Combined Dataset

### 2021-05-04 manuscript/manuscript.tex
revision updates

>>>>

\subsection{Biomass Analysis}

If we have $T$ species, their biomass is $B = \sum_{i=1}^T N_i M_i$ where $N_i$
is their census size and $M_i$ is their wet body mass. In my dataset, I
approximate $N_i$ as the product of a species $i$'s range $R_i$ and their
predicted population density from wet body mass $D_i = \beta_0 + \beta_1 M_i$.
The biomass in my sample, $B'$ is the sum above with some weights $w_i \in \{0,
1\}$; we can write:

\begin{align}
  B' &= \sum_{i=1}^T w_i N_i M_i \\
    &= \sum_{i=1}^T w_i R_i (\beta_0 + \beta_1 M_i ) M_i \\
\end{align}

### 2021-05-04 notebooks/main_analysis.Rmd
revision updates

Based on the predictions of Mora et al. 2011

@@

So, marine species are over represented in my dataset.

@@

# only complete data
xd <- d %>% filter(is.finite(log10_body_mass))

@@

# 1Gt = 1e12 kg
xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * # to kg
                   10^xd$log10_popsize) / 1e12

@@

@@

```{r}
baron_mine <- baron_mine %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)

baron_mine %>% select(phylum, prop_bd_biomass)
```

@@

mamd <- xd %>% filter(class == 'Mammalia') 
mamd$species  <- factor(mamd$species,
                        mamd$species[order(mamd$carbon_mass, decreasing=TRUE)], 
                        ordered=TRUE)
igno <- c("Balaenoptera bonaerensis", "Homo sapiens", "Eschrichtius robustus")
mamd %>% arrange(desc(carbon_mass)) %>% pull(species)

@@

@@

xd %>% filter(species == "Aptenodytes patagonicus") %>% select(species, log10_popsize)

xd %>% filter(genus == "Equus") %>% select(species, log10_popsize, log10_range)

@@

@@

"Homo sapiens"
"Bos taurus"
"Ovis aries"
"Capra hircus"

# using the Bar-On weights
baron_d <- baron_d %>% 
              mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
              mutate(reweighted_total_carbon_mass = reweight * total_carbon_mass)

total_sample_carbon_mass <- baron_d %>% pull(reweighted_total_carbon_mass) %>% sum()

>>>>

@@

Number of species is based on the predictions of Mora et al. 2011. So, marine
species are over represented in my dataset.

This study doesn't have per-class or phylum data, so I build a quick dataset
here from a few sources:

# sources, 
# https://www.environment.gov.au/system/files/pages/2ee3f4a1-f130-465b-9c7a-79373680a067/files/nlsaw-2nd-complete.pdf
nsd <- tribble(~ class, ~ total_species, 
               "Echinoidea", 940,
               "Holothuroidea", 1430,
               "Insecta", 1e6,
               "Aves",  9990,
               "Mammalia", 5487,
               "Clitellata", 8e3, # wikipedia
               "Amphibia", 6515,
               "Demospongiae", 8800, # wikipedia
               "Teleostei", 30e3, # wikipedia
               "Bivalvia", 10e3,  # https://ucmp.berkeley.edu/taxa/inverts/mollusca/bivalvia.php
               "Malacostraca", 40e3, # wikipedia
               "Branchiopoda", 800, # animaldiversity.org
               "Gastropoda", prod(c(40e3, 100e3))^(1/2), # https://ucmp.berkeley.edu/taxa/inverts/mollusca/gastropoda.php
               "Chromadorea", 2124, # ITIS
               "Reptilia", 8734,
               "Ascidiacea", 2300, # wikipedia
               "Anthozoa", 6e3, # wikipedia
               "Euchelicerata", 77000, # wikipedia
               "Holostei", 8, # wikipedia
               "Anopla", 600, # wikipedia
               "Arachnida", 102248,
               "Ophiuroidea",  2064, # https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0031940#
               "Polychaeta", 10000, # https://animaldiversity.org/accounts/Polychaeta/
               "Cephalopoda", 800, # https://ucmp.berkeley.edu/taxa/inverts/mollusca/cephalopoda.php
             "Maxillopoda", 14e3 # https://www.sciencedirect.com/topics/agricultural-and-biological-sciences/maxillopoda
)

Now, compare to the number of species by class/phylum in my data set:

# phyla / class mapping 
phyla_class <- dt_dps %>% select(phylum, class) %>% group_by(phylum, class) %>% 
  distinct()

nsd %>% 
  dt_dps %>% 
  select(phylum, species) %>% group_by(phylum)

# by phyla
dnsp_phyla <- dt_dps %>% 
  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
  select(phylum, species) %>% group_by(phylum) %>% 
               summarize(nspecies=n())

# by class
dnsp_class <- dt_dps %>% 
  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
  select(class, species) %>% group_by(class) %>% 
               summarize(nspecies=n())

dnsp_class %>%
  ggplot(aes(class, log10(frac), size=log10(total_species))) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

By phylum:

dnsp_phyla %>%
  ggplot(aes(phylum, log10(frac), size=log10(total_species))) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

# only complete data
xd <- dt_dps

@@

# 1Gt = 1e12 kg
xd$carbon_mass <- (prop_carbon*(10^xd$log10_body_mass_pred) / 1e3 * # to kg
                   10^xd$log10_popsize) / 1e12

@@

baron_mine %>% left_join(dnsp_phyla, by='phylum')  %>% select(nspecies.x, nspecies.y)
  mutate(a=signif(total_carbon_mass / frac, 2)) %>% 
  select(-bd_total_biomass) %>%
  select(-total_log10_popsize, -nspecies.y) %>% as.data.frame

@@

baron_mine <- baron_mine %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)

baron_mine %>% select(phylum, prop_bd_biomass)

@@

mamd <- xd %>% filter(class == 'Mammalia') 
mamd$species  <- factor(mamd$species,
                        mamd$species[order(mamd$carbon_mass, decreasing=TRUE)], 
                        ordered=TRUE)
#igno <- c("Balaenoptera bonaerensis", "Homo sapiens", "Eschrichtius robustus")
igno <- c()
mamd %>% arrange(desc(carbon_mass)) %>% pull(species)

@@

This seems reasonable... pre-whaling whales, humans, and their domesticates
dominate. Wolves seem high, but this is a macro-ecological based approach that
doesn't reflect hunting to low densities.

Let's look at some random species:

@@

of_interest = c("Aptenodytes patagonicus", "Equus caballus", "Equus ferus
                przewalskii", "Homo sapiens", "Bos taurus", "Ovis aries", "Capra
                hircus", "Canis lupus")

xd %>% filter(species %in% of_interest) %>% select(species, log10_popsize, log10_range)

@@

Ok, now let's look at total biomass. Since I have drastically different phylum
represented here in my taxa sample than on earth, we need to reweight things.

@@

# using the Bar-On weights
# prop_carbon_mass is prop in my dataset
baron_mine <- baron_mine %>% 
              mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
              mutate(reweighted_total_carbon_mass = reweight * total_carbon_mass)

total_sample_carbon_mass <- baron_mine %>% pull(reweighted_total_carbon_mass) %>% sum()
